---
title: "p8105_hw3_sg3891"
author: Sandya Ganesh
date: 2021-10-20
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(p8105.datasets)
```
## Question 1

We will explore the instacart data set

#### Short description of instacart data set

```{r q1_load}
data("instacart")
```

The code below provides the size and structure of the data, describes some key variables, and gives illustrative examples of observations. 

The total number of observations in the instacart data set is `r nrow(instacart)` observations with `r ncol(instacart)` variables. Some key variables in this data set include `r names(instacart %>% select(c(1, 5, 11)))`. There are `r length(unique(pull(instacart, order_id)))` unique orders in this data set, with `r length(unique(pull(instacart, user_id)))` users placing these orders, which means that each user has placed `r round((length(unique(pull(instacart, order_id))))/length(unique(pull(instacart, user_id))), 2)` instacart order in this dataset. These orders include products such as `r pull(instacart, product_name)[1:3]`. This dataset includes orders for `r length(unique(pull(instacart, product_id)))` unique products.

There are `r length(unique(pull(instacart, aisle)))` aisles, and the aisles that the most items are ordered from are `r tail(names(sort(table(pull(instacart,aisle)))), 3)`.

#### Question 1.1

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r q1_plot}

aisle_df = 
  instacart %>% 
  count(aisle) %>% 
  filter(n > 10000)

ggplot(aisle_df, aes(x = reorder(aisle, -n), y = n)) + 
  geom_point() + 
  labs(
    title = "Instacart Popular Aisles ",
    x = "Aisle Names",
    y = "Number of items ordered",
    caption = "Data from the instacart dataset"
  ) +
  coord_flip()
```

#### Question 1.2

Below is a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”, with the number of times each item is ordered.

```{r table1}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  rename(item_count = n) %>% 
  mutate(order_count = min_rank(desc(item_count))) %>% 
  filter(order_count < 4) %>% 
  arrange(aisle, order_count) %>% 
  select(aisle, product_name, item_count) %>% 
  knitr::kable()
```

#### Question 1.3

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r table2, warning = FALSE}
instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name =="Coffee Ice Cream") %>%
  group_by(product_name, order_dow) %>%
  summarize(average = round(mean(order_hour_of_day), 2)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = average
  ) %>%
   rename( "Sunday" = "0","Monday" = "1", "Tuesday" = "2", "Wednesday" = "3", "Thursday" = "4", "Friday" = "5", "Saturday" = "6") %>% 
  knitr::kable()
```

## Question 2

The following question uses the BRFSS data.

#### Clean the BRFSS data
The code chunk below cleans the data:
```{r clean_brfss}
data("brfss_smart2010")

brfss_cleaned = brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic %in% c("Overall Health")) %>%
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) %>%
  arrange(response) %>%
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
  separate(locationdesc, into = c('state', 'location'), sep = ' - ')

```


#### Question 2.1
In 2002, which states were observed at 7 or more locations?
```{r states2002}

brfss_cleaned %>% 
  filter(year == "2002") %>% 
  group_by(state) %>% 
  distinct(location) %>% 
  count(state) %>% 
  filter(n >= 7) %>% 
  select(state)

```
As seen above, the states observed at 7 or more locations include Connecticut, Florida, Massachusetts, North Carolina, New Jersey, and Pennsylvania.

In 2010, which states were observed at 7 or more locations?
```{r states2010}

brfss_cleaned %>% 
  filter(year == "2010") %>% 
  group_by(state) %>% 
  distinct(location) %>% 
  count(state) %>% 
  filter(n >= 7) %>% 
  select(state)

```

As seen above, the states observed at 7 or more locations include California, Colorado, Florida, Massachusetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas, and Washington.


#### Question 2.2
Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 

Additionally, make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r q2_spaghetti}

spaghetti_df = brfss_cleaned %>% 
  filter(response == "Excellent") %>% 
  group_by(state, year) %>% 
  summarize(average = round(mean(data_value, na.rm = TRUE), 2)) %>%
  select(year, state, average)

ggplot(spaghetti_df, aes(x = year, y = average, color = state)) + 
  geom_line() + 
  labs(
    title = "Average value over time for all states ",
    x = "Time (years)",
    y = "Average value",
    caption = "Data from the brfss dataset"
  )

```
#### Question 2.3
Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r q2_panels}

two_2006_2010 = brfss_cleaned %>% 
  filter(year == "2006" | year == "2010") %>% 
  filter(state == "NY")

ggplot(two_2006_2010, aes(x = data_value, color = response)) + 
  geom_density() + 
  facet_grid(. ~ year) +
  labs(
    title = "Distribution of data_value for responses in NY",
    x = "Response",
    y = "data_value",
    caption = "Data from the brfss dataset"
  )

```

## Question 3

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes.

#### Load and clean the accel data

```{r clean_accel}
accel_cleaned = 
  read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>% 
  mutate(
    day_type = ifelse(day == "Saturday" | day == "Sunday", "Weekend", "Weekday"),
    day = factor(day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
    ) %>%
  pivot_longer(
    cols = activity_1:activity_1440,
    names_to = "activity_number",
    values_to = "activity_counts",
    names_prefix = "activity.",
  )
```

Describe the resulting dataset (e.g. what variables exist, how many observations, etc)

Looking into this data set, we can see that the total number of observations in the accel data set is `r nrow(accel_cleaned)` observations with `r ncol(accel_cleaned)` variables. The variables in this data set include `r names(instacart %>% select(c(1,2,3,4,5,6)))`. 

#### Question 3.1
Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r accel_daily}

accel_total_daily = accel_cleaned %>% 
  group_by(day_id) %>% 
  summarize(total_daily_activity = sum(activity_counts)) %>%
  select(day_id, total_daily_activity) %>% 
  knitr::kable()

accel_total_daily
```

Looking at the above table, it appears that the subject had very little activity on day 24 and 31. Other than that, activity appears to fluctuate throughout the 35 days, with no clear trend.

#### Question 3.2

Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.


