---
title: "p8105_hw3_sg3891"
author: Sandya Ganesh
date: 2021-10-20
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(p8105.datasets)
```
## Question 1

We will explore the instacart data set

#### Short Description of instacart data set, with size and structure of the data, describing some key variables, and giving illstrative examples of observations. 
```{r q1_load}
data("instacart")
```

The total number of observations in the instacart data set is `r nrow(instacart)` observations with `r ncol(instacart)` variables. Some key variables in this data set include `r names(instacart %>% select(c(1, 5, 11)))`. There are `r length(unique(pull(instacart, order_id)))` unique orders in this data set, with `r length(unique(pull(instacart, user_id)))` users placing these orders, which means that each user has placed `r round((length(unique(pull(instacart, order_id))))/length(unique(pull(instacart, user_id))), 2)` instacart order in this dataset. These orders include products such as `r pull(instacart, product_name)[1:3]`. This dataset includes orders for `r length(unique(pull(instacart, product_id)))` unique products.

There are `r length((pull(instacart, order_id)))` aisles, and the aisles that the most items are ordered from are `r tail(names(sort(table(pull(instacart,aisle)))), 3)`.



